---
title: Convert R Markdown (Rmd) Files to R Scripts
author: Garrick Aden-Buie
date: '2017-10-17'
slug: convert-r-markdown-rmd-files-to-r-scripts
categories:
  - Blog
tags:
  - pandoc
  - R
  - Scripts
  - workflow
  - Markdown
---

```{r setup}
knitr::opts_chunk$set(cache=TRUE)
```

> **tl;dr:** A [function called backstitch][gist-link] that backstitches `knitr::knit`able R Markdown files to `knitr::spin`able R Scripts.

There are two great ways of writing R Markdown reports.
The most well known way is to use "literate programming", where the writing and the code are intermingled an `.Rmd` file.
This is the style of file you get when you create a new "R Markdown" file inside RStudio.

There's another way of mixing code and writing inside a script, and it's essentially the inverse of an `.Rmd` file.

In an R Markdown file, text is considered to be writing unless you wrap it in decoration, by adding backticks around inline code or by putting code in a "code chunk":

````markdown
`r ''````{r chunk-name}
table(mtcars$cyl, mtcars$gear)
```
````

To get your final report, you call knitr's **`knit`** function, which runs the code chunks, knits the results into the text, and writes an output file (PDF, HTML, Word, etc.).

In the inverse universe, you write your code and text in regular R scripts, mark specific lines as regular text by starting the lines with `#'`, and you **`spin`** your code and text together instead of knitting.

This format is great for computationally-heavy scripts.
The other advantage is that you can call the script directly to run all the code directly, ignoring the text aspect.
I also use it a lot when I want to quickly write up a report on a set of scripts that I've been working with.

Dean Attali has a [great write up from a couple years ago about `spin`][dean-attali] that I highly recommend.
And there's a [quick intro to `spin` in the knitr documentation][yihui-spin], too.

I usually try to keep computation and reporting scripts separate.
But writing up the results of your code while putting it together is an incredible paradigm for data analysis...and sometimes I keep digging and digging while writing and exploring.
And then I end up with a very "heavy" R Markdown (or notebook), where a lot of the data processing code really should be run separately from the reporting and visualization code.

But how do you convert an R Markdown file to a regular R script (or one that you could **`spin`** if you wanted to)?

I couldn't find the answer online, so I wrote a function called `backstitch` that does the conversion.
It only goes one way, for now: from `.Rmd` to `.R`.
But it lets you convert the whole R Markdown document to an R script, or just pull out the code chunks into a script (or both!).

Because it's just a function, I made it [available as a gist][gist-link], which you can source directly with `devtools`.

```{r source-gist, message=FALSE, warning=FALSE}
devtools::source_gist('284671997992aefe295bed34bb53fde6', filename = 'backstitch.R')
```

Give it an input file or connection (`infile`), an output file name (`outfile`, leave as `NULL` to return as text vector), choose `output_type` from `script`, `code`, `both` and it does the rest of the work.
(There's also an extra option called `chunk_header`, for you to set the initial characters for chunk sections.)

For example, using Yihui's [knitr-spin.Rmd](https://github.com/yihui/knitr/blob/master/inst/examples/knitr-spin.Rmd) example, but in reverse, we start with this file:

```{r input, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
inurl <- url('https://raw.githubusercontent.com/yihui/knitr/master/inst/examples/knitr-spin.Rmd')
x <- readLines(inurl)
close(inurl)
cat("~~~r", x, "~~~", sep = '\n')
```

and `backstitch` it into this file:

```{r output, results='asis'}
infile <- url('https://raw.githubusercontent.com/yihui/knitr/master/inst/examples/knitr-spin.Rmd')
output <- backstitch(infile, output_type = 'script', chunk_header = "#+")
cat("```r", output, "```", sep = "\n")
```

[gist-link]: https://gist.github.com/gadenbuie/284671997992aefe295bed34bb53fde6
[dean-attali]: http://deanattali.com/2015/03/24/knitrs-best-hidden-gem-spin/
[yihui-spin]: https://yihui.name/knitr/demo/stitch/