---
title: Style knitr Code Output Appearance in HTML with Custom CSS Classes
description: | 
  Apply custom CSS classes to code chunk outputs to differentiate between 
  regular outputs and messages, warnings, and errors.
author: Garrick Aden-Buie
date: '2019-03-04'
slug: knitr-custom-class-output
categories:
  - Blog
tags:
  - R
  - knitr
  - R Markdown
  - Tips
  - Tricks
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<!-- Links -->
[knitr-og-issue]: https://github.com/yihui/knitr/pull/1335
[knitr-issue]: https://github.com/yihui/knitr/issues/1676
[stackoverflow-question]: https://stackoverflow.com/questions/48538373/chunk-option-class-output-is-not-working-on-error-message
[knitr-example]: https://github.com/yihui/knitr-examples/blob/master/116-html-class.Rmd
[knitr-gh]: https://github.com/yihui/knitr
[knitr-cran]: https://cran.r-project.org/package=knitr

I was working on customizing CSS classes of output chunks to try to achieve the look of code chunks in Brodie Gaslam's blog -- I really like how input and output are visually connected.
Learned about applying custom css classes to code chunks [from Ian Lyttle's pull request][knitr-og-issue] adding support for `class.output`.
But I quickly discovered that this class is only applied to standard code output, and not the "side effect" chunk outputs, such as messages, warnings and errors.

So I opened [an issue in knitr][knitr-issue] and three days later my wishes were granted.
You can now uset the chunk options `class.message`, `class.warning`, and `class.error` to customize the CSS class of outputs of each type.

The new feature will be available in version 1.22 when it makes it to [CRAN][knitr-cran], or now from [GitHub][knitr-gh]

```{r render, include=FALSE}
callr::r(
  function() {
    devtools::dev_mode()
    devtools::install_github("yihui/knitr@cd7435a", dependencies = FALSE)
    rmarkdown::render(here::here("static/blog/2019/knitr-custom-class-output/demo.Rmd"))
    devtools::dev_mode(FALSE)
  }
)
```

<div style="height: 50vh;">
<iframe src="../2019/knitr-custom-class-output/demo.html" width="100%" height="100%">
</iframe>
</div>